Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-05-23 Manuscript/manuscript.Rmd
manuscript edits

We then sought to determine which of the transcription factor binding sites were most influential in predicting TAD boundaries. We found that the cell line specific binding sites associated with SMC3, CTCF, RAD21, ZNF143, and YY1 were given consistently high predictive importance, across all resolutions (Figure 7; Supplementary Figure 6). The SMC3, CTCF, and RAD21 binding sites are known factors associated with the cohesin protein complex, which is hypothesized to be what creates loops in the DNA  [@sanborn2015chromatin; @fudenberg2016formation; @alipour2012self]. Likwise, ZNF143 and YY1 are known structural regulators of enhancer-promoter interactions that also contribute to looping of the linear genome [@weintraub2017yy1; @gonzalez2017znf143]. With the identification of these factors known to be enriched at TAD boundaries, we are able to confirm the validity of our model building process. Additionally, we were also able to identify some lesser known cell line specific transcription factors as predictive of TAD boundaries including the MYC-associated zing-finger protein (MAZ), jun-D proto-oncogene (JUND), and ELF1. These transcription factor binding sites tended to be given higher predictive importance at lower resolution. In fact, at lower resolutions, the predictive importance was found to be much more evenly dispersed across the top transcription factor binding sites, than for higher resolutions. This is again suggestive of the fact that the spatial relationship between functional annotations, specifically transcription factor binding sites, and TAD boundaries changes at different Hi-C data resolutions.

<!--
## DISTANCE PROFILES HIGHLIGHT THE RELATIONSHIP BETWEEN KEY TFBS AND TAD BOUNDARIES

When evaluating the 10 kb data resolution, on GM12878, the distance profiles (on a $log_{2}$ scale) were found to be smaller for top predictive TFBS annotations in relation to genomic bins that overlapped with flanked TAD boundaries, when compared to bins that did not overlap with a flanked TAD boundary (Figure 7). At the same resolution and cell line, direct comparisons of quantiles obtained from the distribution of distances for the transcription factors SMC3a, ZNF143, CTCF, and RAD21, led to the the conclusion that these TFBS were more often located near TAD boundaries on the linear genome, as can be seen from the shifted quantile differences (Figure 7). Similar distance profiles for the same TFBS were obtained when assessed at lower resolutions, across both cell lines (Supplementary Figures 12-17 A-D).
-->

# DISCUSSION

>>>>

We then sought to determine which of the transcription factor binding sites were most influential in predicting TAD boundaries. We found that the cell line specific binding sites associated with SMC3, CTCF, RAD21, ZNF143, and YY1sc281 were given consistently high predictive importance, across all resolutions (Figure 7; Supplementary Figure 6). The SMC3, CTCF, and RAD21 binding sites are known factors associated with the cohesin protein complex, which is hypothesized to be what creates loops in the DNA  [@sanborn2015chromatin; @fudenberg2016formation; @alipour2012self]. Likwise, ZNF143 and YY1 are known structural regulators of enhancer-promoter interactions that also contribute to looping of the linear genome [@weintraub2017yy1; @gonzalez2017znf143]. With the identification of these factors known to be enriched at TAD boundaries, we are able to confirm the validity of our model building process. Additionally, we were also able to identify some lesser known cell line specific transcription factors as predictive of TAD boundaries including the MYC-associated zing-finger protein (MAZ), jun-D proto-oncogene (JUND), and ELF1. These transcription factor binding sites tended to be given higher predictive importance at lower resolution. In fact, at lower resolutions, the predictive importance was found to be much more evenly dispersed across the top transcription factor binding sites, than for higher resolutions. This is again suggestive of the fact that the spatial relationship between functional annotations, specifically transcription factor binding sites, and TAD boundaries changes at different Hi-C data resolutions.

## INFLUENTIAL TFBS TENDED TO BE CLOSER TO BINS THAT OVERLAPPED WITH FLANKED TAD BOUNDARIES THAN THOSE THAT DID NOT

We then investigated the spatial patterns of the most predictive TFBS around TAD boudaries by characterizing the distance profiles between the classes of genomic bins. For GM12878, at 10 kb resolution, we found that SMC3, CTCF, RAD21, ZNF143, and YY1sc281 were all statistically significantly closer to bins that overlapped with flanked TAD boudaries than those that did not (Figure 8). Although still statistically significantly different, the distance profiles between genomic class were found to be much more similar at lower resolutions (Supplementary 9). For the K562 cell line, we reported the distance profiles of the same transcription factor binding sites, excluding the YY1sc281, due to its unavailability in the UCSC Genome Browser. These cell line specific distance profiles were again found to be significantly different between genomic class, at all resolutions (Supplementary 10). 
# DISCUSSION
